---
title: "In Class Exercises - Sorting and Filtering"
author: "Kylie Brownn"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
    toc: true
    toc-location: left
---

## Setup and Loading Data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load the packages we'll need
# Remember, think of this like packing for our trip
# (We don't take the whole closet, only what we need)
library(tidyverse)
library(lubridate)

```

Run this line below load the data for this assignment. It will load a dataframe called "impeach," containing U.S. House Democrats and whether they supported the first impeachment of Pres. Trump, along with when they announced such support.

```{r}

impeach <- read_csv("https://docs.google.com/spreadsheets/d/e/2PACX-1vRh8d5JaDqBtByzNw2JZF3idaACobDhhk-p7chJoktA0HawELHFOvjQOqCpzGA4MGONvPlR7GASqW-K/pub?gid=1765341510&single=true&output=csv")


```

Let's see what we have:

```{r}

impeach


```

## Exercise Questions

*For the questions below, write your R code to return the requested records in the code chunk provided.*

1) The column "for_impeachment" indicates whether the member has publicly called for an impeachment inquiry. Filter to return only the ones where the answer is NO.



```{r}
#filter impeachment that said no
filter(mpeach, for_impeachment == "NO")





```

2) Filter to return only results where a member is both against impeachment, and comes from a district that President Trump won in 2016 (which is noted in the "p16winningparty" column).


```{r}
#filter impeachment that said no and then Trump, who is Republican
mpeach %>%
  filter(for_impeachment == "NO") %>%
  filter(p16winningparty == "R")





```

3) Filter for only results where a member is against impeachment, comes from a district that President Trump won in 2016 (which is noted in the "p16winningparty" column), and also comes from a district that Mitt Romney won in 2012 ("p12winningparty").



```{r}
#filter impeachment answers as no, then Trump (R) and Romney, who is a Republican
mpeach %>%
  filter(for_impeachment == "NO") %>%
  filter(p16winningparty == "R") %>%
  filter(p12winningparty == "R")





```

4) Filter for only results from September 2019 where a member is a YES for impeachment.



```{r}
#filter for Sept month (9) and year (2019) and yes impeachment
mpeach %>%
  filter(date_month == 9) %>%
  filter(date_year == 2019) %>%
  filter(for_impeachment == "YES")




```

5) Filter for only results where a member is a YES for impeachment and is from a district where Clinton won more than 70 percent of the vote in 2016 (found in column "clinton_percent").


  
```{r}
mpeach %>%
  filter(for_impeachment == "YES") %>%
  filter(70 <= clinton_percent)




```

6) Sort the entire dataframe based on the percentage of a district that has a
bachelor's degree or higher ("pct_bachelors"), from lowest to highest.



```{r}
#arrange pct bachelors from lowest to highest
mpeach %>%
  arrange(pct_bachelors)





```

7) Sort the just those who are NO on impeachment based on the percentage of a district that has a bachelor's degree or higher ("pct_bachelors"), from lowest to highest.



```{r}
#filter the impeachment yes answers and then arrange the bachelor degrees
mpeach %>%
  filter(for_impeachment == "YES") %>%
  arrange(pct_bachelors)




```

8) Sort the just those who are NO on impeachment based on the percentage of a district that has a bachelor's degree or higher ("pct_bachelors"), from lowest to highest.

#filter the impeachment no answers and then arrange the bachelor degrees
mpeach %>%
  filter(for_impeachment == "NO") %>%
  arrange(pct_bachelors)
  

Then filter those records by only those whose bachelor's percentage is below the national average (found in the "pct_bachelors_compared_to_national" column).

#same thing as above but filter the averages below national average


```{r}
mpeach %>%
  filter(for_impeachment == "NO") %>%
  arrange(pct_bachelors) %>%
  filter(pct_bachelors_compared_to_national == "BELOW")





```

9) Filter for only members from New Jersey who are NO on impeachment.

```{r}
#filter states of NJ and then filter no impeachment
mpeach %>%
  filter(state == "NJ") %>%
  filter(for_impeachment == "NO")





```

10) Filter for those who were YES on impeachment, with a declared date prior to 2019. So only those with dates before 2019. Then sort those so that the highest Clinton vote percentages are at the top.

```{r}
#filter yes impeachment and years before 2019, then arrange clinton percent from highest to lowest using desc
mpeach %>%
  filter(for_impeachment == "YES") %>%
  filter(date_year <= 2019) %>%
  arrange(desc(clinton_percent))




```
